### 2018-10-02 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

\section{Длабоко учење}

@@

\cite{Sturm2016} A deep neural network is one that has more than one hidden layer of units (neurons) between its input and output layers [25]. Essentially, a neural network transforms an input by a series of cascaded non-linear operations. A recurrent neural network (RNN) is any neural network possessing a directed connection from the output of at least one unit into the input of another unit located at a shallower layer than itself (closer to the input). A deep RNN is a stack of several RNN layers, where each hidden layer generates an output sequence that is then used as a sequential input for the deeper layer. With deeper architectures, one expects each layer of the network to be able to learn higher level representations of the input data and its short- and long-term relationships.
The recurrence (feedback) present in an RNN allows it to take into account
its past inputs together with new inputs. Essentially, an RNN predicts a sequence of symbols given an input sequence. Training it entails modifying the parame- ters of its transformations to diminish its prediction error for a dataset of known sequences. The basic recurrent structure, however, presents problems related to exploding and vanishing gradients during the training procedure [20, 30], which can result in a lack of convergence of solutions. These problems can be circum- vented by defining the hidden layer activation function in a smart way. One such approach defines long short term memory (LSTM) “cells”, which increases the number of parameters to be estimated in training, but controls the flow of information in and out of each cell to greatly help with convergence [16, 21]. Though RNN and LSTM are not new, recent advances in efficient training
algorithms and the prevalence of data have led to great success when they are applied to sequential data processing in many domains, e.g., continuous hand- writing [15], speech recognition [16], and machine translation [35]. In the next subsection, we describe past applications of recurrent networks to music transcription modelling and generation.

@@

\section{Конволуциски мрежи}

\section{Енкодер-Декодер / Авто енкодер}

Encoder-Decoder Architecture for NMT
The Encoder-Decoder architecture with recurrent neural networks has become an effective and standard approach for both neural machine translation (NMT) and sequence-to-sequence (seq2seq) prediction in general.
The key benefits of the approach are the ability to train a single end-to-end model directly on source and target sentences and the ability to handle variable length input and output sequences of text.
As evidence of the success of the method, the architecture is the core of the Google translation service.
Sutskever NMT Model \cite{Sutskever2014}
Deep Neural Networks (DNNs) are powerful models that have achieved excellent performance on difficult learning tasks. Although DNNs work well whenever large labeled training sets are available, they cannot be used to map sequences to sequences.  In this paper, we present a general end-to-end ap proach to sequence learning that makes minimal assumptions on the sequence structure. Our method uses a multilayered Long Short-Term Memory (LSTM) to map the input sequence to a vector of a fixed dimensionality, and then another deep LSTM to decode the target sequence from the vector.  Our main result is that on an English to French translation task from the WMT’14 dataset, the translations produced by the LSTM achieve a BLEU score of 34.8 on the entire test set,  where the LSTM’s BLEU score was penalized on out-of-vocabulary words. Additionally, the LSTM did not have difficulty on long sentences.  For comparison, a phrasebased SMT system achieves a BLEU score of 33.3 on the same dataset.  When we used the LSTM to rerank the 1000 hypotheses produced by the aforementioned SMT system, its BLEU score increases to 36.5, which is close to the previous b est result on this task.  The LSTM also learned sensible phrase and sentence representations that are sensitive to word order and are relatively invariant to the active and the passive voice.  Finally, we found that reversing the order of the words in all source sentences (but not target sentences) improved the LSTM’s performance markedly, because doing so introduced many short term dependencies between the source and the target sentence which made the optimization problem easier.

In this section, we will look at the neural machine translation model developed by Ilya Sutskever, et al. as described in their 2014 paper “Sequence to Sequence Learning with Neural Networks“. We will refer to it as the “Sutskever NMT Model“, for lack of a better name.
This is an important paper as it was one of the first to introduce the Encoder-Decoder model for machine translation and more generally sequence-to-sequence learning.
The model was applied to English to French translation, specifically the WMT 2014 translation task.
The translation task was processed one sentence at a time, and an end-of-sequence (<EOS>) token was added to the end of output sequences during training to signify the end of the translated sequence. This allowed the model to be capable of predicting variable length output sequences.
    Note that we require that each sentence ends with a special end-of-sentence symbol “<EOS>”, which enables the model to define a distribution over sequences of all possible lengths.
The model was trained on a subset of the 12 Million sentences in the dataset, comprised of 348 Million French words and 304 Million English words. This set was chosen because it was pre-tokenized.
The source vocabulary was reduced to the 160,000 most frequent source English words and 80,000 of the most frequent target French words. All out-of-vocabulary words were replaced with the “UNK” token.
Model
An Encoder-Decoder architecture was developed where an input sequence was read in entirety and encoded to a fixed-length internal representation.
A decoder network then used this internal representation to output words until the end of sequence token was reached. LSTM networks were used for both the encoder and decoder.
    The idea is to use one LSTM to read the input sequence, one timestep at a time, to obtain large fixed-dimensional vector representation, and then to use another LSTM to extract the output sequence from that vector
The final model was an ensemble of 5 deep learning models. A left-to-right beam search was used during the inference of the translations.

Cho NMT Model \cite{Cho2014}
In this section, we will look at the neural machine translation system described by Kyunghyun Cho, et al. in their 2014 paper titled “Learning Phrase Representations using RNN Encoder–Decoder for Statistical Machine Translation.” We will refer to it as the “Cho NMT Model” model for lack of a better name.
Importantly, the Cho Model is used only to score candidate translations and is not used directly for translation like the Sutskever model above. Although extensions to the work to better diagnose and improve the model do use it directly and alone for translation.
he model uses the same two-model approach, here giving it the explicit name of the encoder-decoder architecture. called RNN Encoder–Decoder that consists of two recurrent neural networks (RNN). One RNN encodes a sequence of symbols into a fixed-length vector representation, and the other decodes the representation into another sequence of symbols. 
In  this  paper,  we  propose  a  novel  neural network model called RNN Encoder–Decoder  that  consists  of  two  recurrent neural  networks  (RNN).  One  RNN  encodes a sequence of symbols into a fixed length vector representation, and the other decodes the representation into another sequence of symbols.  The encoder and decoder  of  the  proposed  model  are  jointly trained to maximize the conditional probability of a target sequence given a source sequence.   The performance of a statistical machine translation system is empirically found to improve by using the conditional probabilities of phrase pairs computed by the RNN Encoder–Decoder as an additional feature in the existing log-linear model.    Qualitatively,  we  show  that  the proposed model learns a semantically and syntactically meaningful representation of linguistic phrases

>>>>

\subsection{Длабоко учење}

@@

Длабоките невронски мрежи се изградени од повеќе слоја од невронски ќелии (или јазли), од кои барем два се задолжителни, а тоа се влезниот и излезлниот слој. Влезниот слој ги претставува податоците на нервонската мрежа а излезниот слој ја врши задачата на мрежа, избор на класа ако е класификација, пресметка на регресиона вредност и тн. Помеѓу овие два слоја мрежата има еден или повеќе скриени слоеви, коишто го вршат адаптирањето на атрибутите кон податочното множество. (СЛИКА ОД ОСНОВНА НЕВРОНСКА МРЕЖА) Секоја ќелија во мрежата има барем еден влез и еден излез, поточно ќелиите од скриениот слој има повеќе влеза и излеза, ќелиите од влезниот слој обично само по еден влез а повеќе излеза, а ќелиите од излезниот слој обично имаат повеќе влеза и еден излезен слој. (СЛИКА ОД ЌЕЛИЈА) Излезот од ќелијата претставува математичка функција од влезовите модифицрани со тежински параметар (ФОРМУЛА ЗА ЌЕЛИЈА). При процесот на учење се модифицираат тежинските параметри на мрежата.
Тековно најчесто користени типови на длабоки невронски мрежи се:
\begin{itemize}
    \item Feed-Forward Neural Networks - Наједноставен тип на невронски мрежи, составени од елементарни невронски ќелии. Работат добро со податоци со многу шум и се едноставни за оддржување. Се користат често во системи за компјутерска визија и препознавање на глас како и како градбени компоненти во посложени мрежи.
    \item Convolutional Neural Networks - Слични на Feed-Forward мрежите, имаат пирамидална структура, каде што секој повисок слој е задолжен за вадење на помала апстрактна репрезентација на дел од слојот под него. Многу често се користи во компјутерска визија и обработка/класификација на слики и сигнали.
    \item Recurrent Neural Networks - Слични на Feed-Forward мрежите со тоа што секоја ќелија има сопствена меморија, т.е. има пристап до своите излези во последните N чекори, каде N е должина на меморијата на мрежата. Запомнетите излези доаѓаат на влез на ќелијата и се дел од пресметките на новиот излез. Најпопуларни варијанти се Long Short Term Memory и Gater Recurrent Unit. Најчесто се користат при работа со секвенцијални податоци, како на пример предвидување на временски серии, машински превод, препознавање на глас, препознавање на ракопис и тн.
    \item Хибридни мрежи се составени од повеќе видови на слоеви од невронски мрежи или цели подмрежи, и секој од деловите на мрежата придонесува со својата под-задача кон главната задача. Подмрежите обично немаат споделени врски помеѓу сопствените скриени слоеви и слоевите на другите подмрежи.
\end{itemize}

@@
